---
title: "Branding and Sidebar Modes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Branding and Sidebar Modes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(bs4Dashkit)
```

## `dash_titles()` — full argument reference

`dash_titles()` is the single constructor for all branding and sidebar behaviour.
It returns a list you destructure into the three placement slots described below.

```{r}
ttl <- dash_titles(
  brand_text     = "My App",       # primary navbar label and default expanded sidebar label
  app_name       = NULL,            # browser tab title; defaults to brand_text
  icon           = icon("shield-halved"), # or a Font Awesome name like "shield-halved"
  icon_img       = NULL,            # image path — overrides icon when set
  icon_shape     = "circle",        # "circle" | "rounded" | "square"
  icon_size      = NULL,            # CSS size e.g. "22px"
  icon_color     = NULL,            # CSS color for FA icon e.g. "#2f6f8f"
  weight         = 700,             # CSS font-weight
  spacing        = "-0.02em",       # CSS letter-spacing
  size           = NULL,            # CSS font-size for label e.g. "13px"
  italic         = FALSE,
  font_family    = NULL,            # CSS font-family string
  color          = NULL,            # solid label colour; ignored when gradient set
  gradient       = NULL,            # c("#hex1", "#hex2") — sets gradient effect
  effect         = "none",          # "none" | "glow" | "shimmer" | "emboss"
  glow_color     = NULL,            # colour for glow / shimmer
  collapsed      = "icon-only",     # brand mode when sidebar is collapsed
  expanded       = "icon-text",     # brand mode when sidebar is expanded
  collapsed_text = NULL,            # optional short label for collapsed icon-text/text-only
  expanded_text  = NULL             # optional expanded override; defaults to brand_text
)
```

### Return value

| Slot | Type | Placement |
|---|---|---|
| `ttl$app_name` | `character` | `bs4DashPage(title = ttl$app_name)` |
| `ttl$brand` | `tagList` | `bs4DashNavbar(title = ttl$brand)`; the sidebar brand mirrors this automatically in a standard `bs4DashPage()` |
| `ttl$deps` | `tagList` | inside `bs4DashBody(...)` — handled by `use_bs4Dashkit_core()` |

---

## Icon options

### Font Awesome icon

```{r}
ttl <- dash_titles(
  brand_text = "My App",
  icon       = icon("chart-line"),
  icon_color = "#2f6f8f",
  icon_size  = "20px"
)
```

### Image logo

`icon_img` overrides `icon`. The path is relative to your app's `www/` folder.

```{r}
ttl <- dash_titles(
  brand_text = "My App",
  icon_img   = "logo.png",    # served from www/logo.png
  icon_shape = "rounded",     # "circle" | "rounded" | "square"
  icon_size  = "28px"
)
```

---

## Reusable brand options

Use `dash_brand_options()` when multiple apps share a brand but each app still
needs the regular `dash_titles()` output:

```{r}
brand <- dash_brand_options(
  name      = "Atlas Forge",
  icon      = icon("layer-group"),
  collapsed = "AF",
  expanded  = "Atlas Forge"
)

ttl <- dash_titles_from(brand)
```

Any extra arguments are forwarded to `dash_titles()`, and direct arguments to
`dash_titles_from()` override stored options:

```{r}
brand <- dash_brand_options(
  name = "Atlas Forge",
  icon = icon("layer-group"),
  effect = "glow",
  glow_color = "#2f6f8f"
)

ttl <- dash_titles_from(brand, app_name = "Atlas Forge Ops")
```

---

## Label effects

```{r}
# No effect (plain text)
ttl <- dash_titles(brand_text = "My App", effect = "none")

# Glow
ttl <- dash_titles(brand_text = "My App", effect = "glow", glow_color = "#2f6f8f")

# Shimmer (animated gradient sweep)
ttl <- dash_titles(brand_text = "My App", effect = "shimmer", glow_color = "#5ba3c9")

# Emboss (subtle relief shadow)
ttl <- dash_titles(brand_text = "My App", effect = "emboss")

# Two-colour gradient (effect argument is ignored when gradient is set)
ttl <- dash_titles(brand_text = "My App", gradient = c("#2f6f8f", "#5ba3c9"))
```

---

## Sidebar modes

The sidebar brand area can show different content depending on whether the sidebar
is collapsed or expanded. There are three modes, combinable independently for each state.

| Mode | What renders |
|---|---|
| `"icon-only"` | Icon or image only — no text |
| `"icon-text"` | Icon plus a text label |
| `"text-only"` | Text label only — no icon |

`brand_text` is the default expanded sidebar label. Reach for
`collapsed_text` when the narrow sidebar needs a shorter label, and use
`expanded_text` only when the expanded sidebar should differ from `brand_text`.
In practice, collapsed labels work best at about 3 characters or fewer.
When `collapsed = "text-only"`, the text is centered in the collapsed brand strip.

### Common combinations

```{r}
# Compact when collapsed, full when expanded (recommended default)
ttl <- dash_titles(
  brand_text     = "Test Dashboard",
  icon           = icon("project-diagram"),
  collapsed      = "icon-only",
  expanded       = "icon-text"
)

# Show abbreviated text when collapsed
ttl <- dash_titles(
  brand_text     = "OLTCR Dashboards",
  icon           = icon("project-diagram"),
  collapsed      = "icon-text",
  expanded       = "icon-text",
  collapsed_text = "OLT",            # about 3 chars works best
  expanded_text  = "OLTCR Dashboards"
)

# Icon only in both states
ttl <- dash_titles(
  brand_text = "OLTCR Dashboards",
  icon       = icon("project-diagram"),
  collapsed  = "icon-only",
  expanded   = "icon-only"
)

# Textless icon brand in both states
ttl <- dash_titles(
  brand_text = NULL,
  app_name   = "Icon Lab",
  icon       = icon("cloud"),
  collapsed  = "icon-only",
  expanded   = "icon-only"
)
```

---

## Hover-expand sidebar

`use_bs4Dashkit_core()` includes hover-expand sidebar behavior by default.
When the sidebar is collapsed on desktop, it expands temporarily on mouse hover.

Configure it with the sizing arguments:

```{r}
body = bs4DashBody(
  use_bs4Dashkit_core(
    ttl,
    topbar_h = "56px",
    collapsed_w = "4.25rem",
    expanded_w = "270px"
  ),
  # ...
)
```

Numeric values remain supported for compatibility: `topbar_h` uses pixels,
`collapsed_w` uses rem, and `expanded_w` uses pixels. Explicit CSS units are
recommended in new code because they avoid guessing which unit a number means.

If you are not using use_bs4Dashkit_core(), you can enable hover-expand directly:

```{r}
body = bs4DashBody(
  use_bs4Dashkit(),
  ttl$deps,
  use_dash_sidebar_behavior(),
  # ...
)
```
---

## Brand divider

A subtle horizontal rule between the brand mark and the sidebar menu:

```{r}
ttl <- dash_titles(
  brand_text    = "My App",
  icon          = icon("chart-line"),
  brand_divider = TRUE           # default TRUE (can be disabled)
)

# Or control it independently
use_dash_sidebar_brand_divider()
```

---

## Global options

Set defaults for your whole project in `global.R` or `app.R` before calling
`dash_titles()`:

```{r}
options(
  bs4Dashkit.sidebar.collapsed = "icon-only",
  bs4Dashkit.sidebar.expanded  = "icon-text",
  bs4Dashkit.brand_divider     = TRUE
)
```

`dash_titles()` reads these options when the corresponding arguments are `NULL`,
so individual apps only need to specify what differs from the project default.

---

## Putting it together

```{r}
library(shiny)
library(bs4Dash)
library(bs4Dashkit)

ttl <- dash_titles(
  brand_text     = "Test Dashboard",
  icon           = icon("project-diagram"),
  weight         = 700,
  effect         = "shimmer",
  glow_color     = "#2f6f8f",
  size           = "20px",
  collapsed      = "icon-text",
  expanded       = "icon-text",
  collapsed_text = "DT",
  expanded_text  = "Dashboard",
  brand_divider  = TRUE
)

ui <- bs4DashPage(
  title  = ttl$app_name,
  header = bs4DashNavbar(title = ttl$brand, skin = "light"),
  sidebar = bs4DashSidebar(
    skin = "light",
    bs4SidebarMenu(
      bs4SidebarMenuItem("Dashboard", tabName = "dash", icon = icon("gauge-high")),
      bs4SidebarMenuItem("Projects",  tabName = "proj", icon = icon("folder-open"))
    )
  ),
  body = bs4DashBody(
    bs4TabItems(
      bs4TabItem(tabName = "dash", h3("Dashboard")),
      bs4TabItem(tabName = "proj", h3("Projects"))
    )
  )
)

server <- function(input, output, session) {}
shinyApp(ui, server)
```

