---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(bs4Dashkit)
```

## Overview

`bs4Dashkit` extends `{bs4Dash}` by standardizing branding, sidebar behavior,
navbar utilities, and theme customization. Every feature is optional — you can
adopt as little or as much as you need.

The package is organised around four areas:

1. **Brand and sidebar** — `dash_titles()` + `use_bs4Dashkit_core()`
2. **Theming** — `use_dash_theme()` and named presets
3. **Navbar** — `dash_nav_title()`, status / refresh / help items, validation, and user menu
4. **Footer** — `dash_footer()` with flexible logo and text positioning

---

## Minimal example

The smallest possible `bs4Dashkit` app requires exactly two calls beyond normal
`bs4Dash` boilerplate:

```{r}
library(shiny)
library(bs4Dash)
library(bs4Dashkit)

# 1. Build the brand object
ttl <- dash_titles("My Dashboard", icon = icon("cloud"))

ui <- bs4DashPage(
  title   = ttl$app_name,           # browser tab title
  header  = bs4DashNavbar(title = ttl$brand),
  sidebar = bs4DashSidebar(),
  body    = bs4DashBody(
    use_bs4Dashkit_core(ttl)        # 2. always the first call in body and called once
  )
)

server <- function(input, output, session) {}
shinyApp(ui, server)
```

`dash_titles()` accepts a single string and sensible defaults handle everything
else. When no icon is supplied, the sidebar brand defaults to text-only modes.
`use_bs4Dashkit_core()` loads the CSS dependencies, activates the sidebar
brand mode, and applies the default theme.

If you prefer, you can also pass a simple `shiny::icon()` tag:

```{r}
ttl <- dash_titles(
  brand_text = "My Dashboard",
  icon = icon("cloud")
)
```

---

## `dash_titles()` — the brand object

`dash_titles()` is the single entry point for all brand-related configuration.
It returns a named list with three slots:

| Slot | Where it goes |
|---|---|
| `ttl$app_name` | `bs4DashPage(title = ...)` |
| `ttl$brand` | `bs4DashNavbar(title = ...)`; the sidebar brand mirrors this automatically in a standard `bs4DashPage()` |
| `ttl$deps` | inside `bs4DashBody(...)` (handled automatically by `use_bs4Dashkit_core()`) |

```{r}
ttl <- dash_titles(
  brand_text     = "OLTCR Dashboards",
  icon           = icon("project-diagram"),
  weight         = 700,
  effect         = "shimmer",          # "none" | "glow" | "shimmer" | "emboss"
  glow_color     = "#2f6f8f",
  size           = "20px",
  collapsed      = "icon-text",
  expanded       = "icon-text",
  collapsed_text = "OLT",              # shown when sidebar is collapsed
  expanded_text  = "OLTCR Dashboards",
  brand_divider  = TRUE
)
```

`brand_text` is the primary label for the app. It is always used for the navbar
brand and is also the default expanded sidebar label. Use `expanded_text` only
when the expanded sidebar should say something different, and use
`collapsed_text` when the collapsed sidebar needs a shorter label. In practice,
about 3 characters works best in the narrow collapsed sidebar.

If you want a fully textless icon brand, set `brand_text = NULL` and keep both
sidebar states at `"icon-only"`:

```{r}
ttl <- dash_titles(
  brand_text = NULL,
  app_name   = "Icon Lab",
  icon       = icon("cloud"),
  collapsed  = "icon-only",
  expanded   = "icon-only"
)
```

For app suites that share the same brand, you can keep the brand options in a
small reusable object:

```{r}
brand <- dash_brand_options(
  name      = "Atlas Forge",
  icon      = icon("layer-group"),
  collapsed = "AF",
  expanded  = "Atlas Forge"
)

ttl <- dash_titles_from(brand)
```

Alternatively, use an image instead of a Font Awesome icon:

```{r}
ttl <- dash_titles(
  brand_text = "My App",
  icon_img   = "logo.png",        # overrides `icon`
  icon_shape = "rounded"              # "circle" | "rounded" | "square"
)
```

**NB: Place logo.png in your app’s www/ directory.**

---

## `use_bs4Dashkit_core()` — the core entry point

Always place this as the **first element** inside `bs4DashBody()`. It:

- Injects the package CSS and JavaScript
- Activates the sidebar brand mode logic
- Applies a theme preset (default: `"professional"`)

```{r}
body = bs4DashBody(
  use_bs4Dashkit_core(ttl, preset = "professional"),
  # ... rest of your content
)
```

Available presets:

```{r}
use_bs4Dashkit_core(ttl, preset = "professional")  # cool blue-grey (default)
use_bs4Dashkit_core(ttl, preset = "modern")        # brighter accent colours
use_bs4Dashkit_core(ttl, preset = "dark-lite")     # dark surfaces and lighter text
bs4dashkit_theme_presets()                         # list built-in presets
```

For a runnable template app, use:

```{r}
app <- bs4dashkit_example_app()
```

For a fuller interactive playground that exercises sidebar states, hover
behavior, presets, navbar controls, and footer styling together, use:

```{r}
app <- bs4dashkit_demo_app()
```

You can also launch the packaged example directory directly:

```{r}
shiny::runApp(system.file("examples", "real-shiny-app", package = "bs4Dashkit"))
```

That packaged example contains the full app code directly in
`inst/examples/real-shiny-app/app.R`, so it can be read, modified, and run as
a standalone reference app.

There is also a heavier shipped example that exercises the navbar title
alignment controls, sidebar text sizing and weights, icon-only branding, and
theme presets together:

```{r}
shiny::runApp(system.file("examples", "test-all", package = "bs4Dashkit"))
```

---

## Sidebar modes

The sidebar brand area can show different content depending on whether the sidebar
is collapsed or expanded. Set these in `dash_titles()`:

```{r}
ttl <- dash_titles(
  brand_text     = "OLTCR Dashboards",
  icon           = icon("chart-line"),
  collapsed      = "icon-only",       # just the icon when narrow
  expanded       = "icon-text",       # icon + label when wide
  collapsed_text = "OLT",
  expanded_text  = "OLTCR Dashboards" # optional; brand_text is the default
)
```

| Mode | Collapsed | Expanded |
|---|---|---|
| `"icon-only"` | ✓ | ✓ |
| `"icon-text"` | ✓ | ✓ |
| `"text-only"` | ✓ | ✓ |

Add hover-expand behavior (sidebar expands on mouse-over even when collapsed):

```{r}
body = bs4DashBody(
  use_bs4Dashkit_core(
    ttl,
    preset = "professional",
    topbar_h = "56px",
    collapsed_w = "4.25rem",
    expanded_w = "270px"
  )
  # ...
)
```
**NB: `use_dash_sidebar_behavior()` is included by `use_bs4Dashkit_core()`.** Call it directly only if you are not using the core helper.

Hover-expand behavior is included by `use_bs4Dashkit_core()`. Configure it via `topbar_h`, `collapsed_w`, and `expanded_w`. Numeric values remain supported with the historical units (`px`, `rem`, `px` respectively), but explicit CSS units are clearer and recommended. Call `use_dash_sidebar_behavior()` directly only when not using the core helper.

---

## Navbar structure

For `bs4DashNavbar(rightUi = ...)`, prefer the complete item helpers. They
return the `<li class="nav-item dropdown">` wrapper expected by bs4Dash:

```{r}
rightUi <- tagList(
  dash_nav_status_item("Ready", status = "success", icon = icon("circle-check")),
  dash_nav_refresh_item("refresh"),
  dash_nav_help_item("help")
)

validate_bs4dash_navbar(rightUi)
```

Use `dash_nav_item()` for custom controls that do not already return a navbar
`<li>`. `validate_bs4dash_navbar()` is useful when mixing custom and packaged
navbar components.

---

## Global options

Set package-level defaults once (e.g. in your `global.R`) so every app in a
project inherits them:

```{r}
options(
  bs4Dashkit.sidebar.collapsed = "icon-only",
  bs4Dashkit.sidebar.expanded  = "icon-text",
  bs4Dashkit.brand_divider     = TRUE,
  bs4Dashkit.accent            = "#2f6f8f",
  bs4Dashkit.theme_preset      = "professional"
)
```

If `preset = NULL`, `use_bs4Dashkit_core()` uses the package option `bs4Dashkit.theme_preset (if set)`. If `accent = NULL`, it uses `bs4Dashkit.accent`.

```{r}
use_bs4Dashkit_core(ttl, preset = NULL) # uses option if set
use_bs4Dashkit_core(ttl, accent = NULL) # uses option if set
```

`dash_titles()` reads these options when the corresponding arguments are `NULL`,
so you can omit them from individual calls.

---

## Troubleshooting

- "I see 404s for dash-theme.css" -> call use_bs4Dashkit_core() or use_bs4Dashkit() first (resource path)

- "My icon doesn’t show" -> check Font Awesome name / include icon_img

- "Hover expand doesn’t work on mobile "-> expected (touch devices)

- "Sidebar labels do not match the hover-expanded state" -> update to the current package version so the shipped hover-expanded sidebar rules are used


## What next?

- **Branding and sidebar modes** — full reference for all `dash_titles()` arguments
- **Theming and presets** — CSS variable overrides and custom presets
- **Navigation utilities** — status, refresh, help, user menu, validation, and `dash_nav_title()`
- **Footer** — all footer layout combinations
- **Complete example app** — a single app that exercises every feature



