---
title: "Packaged Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Packaged Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

`bs4Dashkit` ships with three runnable example apps under `inst/examples/`.
These are useful for smoke testing, regression checks, and quick reference when
you want to compare package behavior against a known working app.

## Included examples

### Real Shiny App

The main interactive playground for sidebar brand modes, navbar title
alignment, theme presets, footer styling, and live brand controls.

```{r}
shiny::runApp(system.file("examples", "real-shiny-app", package = "bs4Dashkit"))
```

Source file:

```{r}
file.show(system.file("examples", "real-shiny-app", "app.R", package = "bs4Dashkit"))
```

### Sidebar Smoke Test

A compact app focused on the sidebar brand behavior. Use this when you want to
quickly verify collapsed labels, hover-expanded labels, and stable sidebar menu
item positioning.

```{r}
shiny::runApp(system.file("examples", "sidebar-smoke-test", package = "bs4Dashkit"))
```

Source file:

```{r}
file.show(system.file("examples", "sidebar-smoke-test", "app.R", package = "bs4Dashkit"))
```

### Test All

A heavier end-to-end example that exercises the package more aggressively:
centered navbar titles, collapsed text-only branding, icon-only branding,
collapsed and expanded text size / weight controls, theme presets, alerts,
badges, and footer styling.

```{r}
shiny::runApp(system.file("examples", "test-all", package = "bs4Dashkit"))
```

Source file:

```{r}
file.show(system.file("examples", "test-all", "app.R", package = "bs4Dashkit"))
```

## When to use which

- Use `real-shiny-app` when you want the interactive demo.
- Use `sidebar-smoke-test` when you are debugging sidebar regressions.
- Use `test-all` when you want a broader package-level stress test.
